Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

### 2022-02-05 GitControlled/Writing/Re_JournalSubmission/ML_VRA_v2.rmd
Large revise of the manuscript and codes for re-submission

---
title: "Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation"
author: Shunkei Kakimoto^[Corresponding author, Department of Agricultural Economics, University of Nebraska-Lincoln 102 Filley Hall 1625 Arbor Drive Lincoln, NE 68583, USA, E-mail address; skakimoto3@hunskers.unl.edu], Taro Mieno^[Department of Agricultural Economics, University of Nebraska-Lincoln, Lincoln, NE 68583, USA], Takashi S. T. Tanaka^[Artificial Intelligence Advanced Research Center, Faculty of Applied Biological Sciences, Gifu University, Gifu 501-1193, Japan], and David S. Bullock^[Department of Agricultural and Consumer Economics, University of Illinois, Urbana, IL 61801, USA]
output:
  bookdown::word_document2:
    fig_caption: yes
    number_sections: yes
    global_numbering: true
    # reference_docx: word_template.docx
    reference_docx: word_template_format.docx
    # pandoc_args: ["-Fpandoc-crossref"]
bibliography: ML_VRA.bib
csl: computers-and-electronics-in-agriculture.csl
eqnPrefix: 
      - "Eq"
abstract: "Estimating site-specific crop yield response to changes in input (e.g., seed, fertilizer) management is a critical step in making economically optimal site-specific input recommendations. Past studies have attempted to estimate yield response functions using various Machine Learning (ML) methods, including the Random Forest (RF), Boosted Random Forest (BRF), and Convolutional Neural Network (CNN) methods. This study proposes use of the Causal Forest (CF) model, which is one of the emerging ML methods that comprise \"Causal Machine Learning.\" Unlike previous prediction-oriented ML methods, CF focuses strictly on estimating heterogeneous treatment effects of inputs, rather than predicting yield. We report results of using Monte Carlo simulations assuming various production scenarios to test the effectiveness of CF in estimating site-specific economically optimal nitrogen rates (EONRs), comparing CF with the yield-prediction-oriented ML methods RF, BRF, and CNN. CF's estimations of site-specific EONRs were superior under all scenarios considered. Furthermore, we also show that the quality of a model's yield predictions provides little if any information about the quality of its EONR predictions."  
---

@@

# opts_knit$set(root.dir = here())
opts_knit$set(root.dir = "~/Dropbox/ResearchProject/ML_VRA")

@@

```{r, cache = FALSE}
# === packages ===#
library(sf)
library(png)
library(data.table)
library(grf)
library(RColorBrewer)
library(patchwork)
library(plotly)
library(magrittr)
library(ggplot2)
library(viridis)
library(tmap)
library(tidyverse)
library(ggthemes)
library(mgcv)
library(ggpubr)
library(flextable)
library(officer)
library(officedown)
library(modelsummary)
library(raster)
library(bookdown)
library(png)
library(grid)
```

```{r analyses, results = "hide", fig.show='hide', cache = FALSE}
#/*=================================================*/
#' # Preparation
#/*=================================================*/
# This is where you prepare data sets and results 
# library(here)
# setwd(here())
#/*----------------------------------*/
#' ## Create figures and tablesb
#/*----------------------------------*/
# --- Convert an Rmd file into a pure R file using purl() ---#
knitr::purl("./GitControlled/Codes/PrepareResults.Rmd", output = "./GitControlled/Codes/PrepareResults.R")

# --- Read the simulation results and prepare figures and tables ---#
source("./GitControlled/Codes/PrepareResults.R")
```

@@

Despite the potential economic and environmental benefits of site-specific input (e.g., nitrogen (N), seed) management, the adoption rate of variable rate application technology remains relatively low [@lowenberg2019setting]. 
This is primarily because of the difficulty in estimating site-specific yield response functions accurately within a field, which is critical for the identification of site-specific economically optimal input rates (EOIRs) [@bullock1998does].

Historically, site-specific yield response functions have been estimated using the data generated from strip trials conducted on only parts of fields [@makowski2001statistical; @scharf2002corn; @bullock2002adding; @mamo2003spatial; @hurley2004estimating; @anselin2004spatial; @scharf2005field; @ruffo2006site; @Miao2006; @puntel2019development].
But recently whole-field on-farm precision experimentation (OFPE) [@bullock2019] has allowed implementation of improved statistical experimental designs (e.g., Latin Square, Randomized Block) are used for an entire field [@laurent_kyveryga_makowski_miguez_2019; @licht_witt_2019]. Data generated from whole-field OFPEs typically have greater local variation in experimental input application rates and soil/field characteristics than do strip trial designs, which has improved the statistical identification of heterogeneous impacts of input management changes on yields. The resultant availability of more spatially detailed experiment data has encouraged a resurgence of interest in estimating site-specific yield response functions. For example, @barbosa2020modeling, @krause2020random and @Gardner21 used OFPE data to estimate site-specific yield response functions and economically optimal site-specific N fertilizer and/or seed rates. These recent studies used machine learning (ML) methods extensively in place of more traditional parametric modeling approaches. @barbosa2020modeling applied the Convolutional Neural Network approach, @krause2020random used Random Forest-based approaches, and @Gardner21 used the shape-constrained generalized additive model. But this trend of applying ML methods may raise concerns. Historically, agronomy is a field that has paid meticulous attention to identifying the "causal" impacts of input management on yield through clever experimental designs. But the aforementioned recent applications of ML methodologies pay little attention to causal identification of inputs on yield, instead validating their models by yield prediction accuracy. But the ultimate goal of OFPE is the accurate prediction of EOIRs, not yields. These are distinct objectives, and achievement of one need not imply the other.

The lack of and need for causal interpretation of the results of ML research applications have been increasingly recognized, in particular in economics and medicine [@athey2017beyond; @athey2018impact; @Athey2019a; @moraffah2020causal; @Storm2020; @arti2020research; @scholkopf2021toward]. In response, causal machine learning (CML) methods were recently developed. Unlike traditional prediction-oriented ML methods, CML focuses on identifying causal impacts of an event (in our context, an increase or decrease in the input application rate). The growing academic literature applying CML methods includes @richens2020improving, who pointed out that the inability of the traditional ML methods to identify cause and effect relationships between disease and patients' characteristics could lead to misdiagnoses, and showed that CML-based diagnoses could improve the accuracy of diagnostic prediction. In addition, @carbo2020machine used a CML to identify characteristics causing bank customers to adopt online banking services, and @bozorgi2020process applied CML to study the mining processes.

In this paper, we test the accuracy of site-specific economically optimal N fertilization rates (EONRs) estimation based on Causal Forest (CF) approach developed by @athey2016recursive and @Wager2018a. We compare the EONR prediction performances of our CF-based approach to the widely used ML methods Random Forest (RF), Boosted Random Forest (BRF), and Convolutional Neural Network (CNN). Further, we also test how yield prediction performance is related to EONR prediction performance to examine whether it is appropriate to select the model based on its ability to predict yield when the ultimate goal is to identify EOIR. We use Monte Carlo simulations, where we synthetically generate yield and soil/field characteristics datasets for a stylized "field." Because real-world "true" EOIRs cannot be measured, real-world data cannot be used to test the accuracy of statistical methods estimating EOIRs. This important difference contrasts starkly with yield prediction, since actual yields can be ground-truthed.

<!-- Our simulation results reveal that CF-based approaches provide superior performance in site-specific EOIR estimation. We also demonstrate that a method's yeidl prediction accuracy is not a good indicator of its EOIR estimation accuracy. -->

# Predicting Yield vs. Predicting Changes in Yield in Response to Changes in Input Management

In the following, we describe the fundamental information necessary to identify EOIR in general: the site-specific treatment effect of input application on yield. We then discuss the causal forest method in detail, show how it is particularly suitable to obtain this fundamental information, and contrast it to more traditional prediction-oriented ML methods, such as RF, BRF, and CNN.

@@

Elementary calculus shows that the necessary condition for maximization is, that $\partial f(I, \mathbf{c})/\partial I - w =0$, which implicitly defines the economically optimal input application rate as a function of prices and soil/field characteristics, and may be arranged as \@ref(eq:eqn2):

@@

Identity \@ref(eq:eqn2) states that the necessary elements of the information needed for profit maximization are the input price, the output price, and the partial derivative of the yield function with respect to the input. While knowledge of the yield function $f(I, \mathbf{c})$ in Eq. \@ref(eq:eqn1)) is sufficient to solve the maximization problem above (since its derivative can be calculated), identity \@ref(eq:eqn2) shows that it is not a necessary condition; rather, knowing the function's partial derivative is necessary.  In the following, we discuss a method of using Causal Forest methods that can provide statistical advantages to the estimation of economically optimal input application rates because it is built to estimate approximations of the derivative $\partial f/\partial I$ directly, and not the yield function $f$ itself.

@@

Recent papers using ML methods to estimate the effectiveness of various agricultural input management strategies have based their analyses on estimates of a yield function, $f(I, \mathbf{c})$ (e.g., @krause2020random; @barbosa2020modeling; @Gardner21). In the following, we present a Causal Forest methodology that does not rely on estimation of how the level of yield depends on the input application rate $I$, but rather relies on a discrete version of Eq. \@ref(eq:eqn2), examining the discrete changes in yield that result from discrete *changes* in the input application rate $I$.

Consider an increasing ordering of input application rates $I_1, I_2,...,I_K$. Among these K rates, for a generic site $i$ a specific EOIR can be identified if the *treatment effect* function $\Delta y_i(I_{k+1}, I_{k})$ is known. This function is defined as the discrete change in yield that results from a discrete change in the input level from $I_k$ to $I_{k+1}$, and it depends on the characteristics at the site analyzed:

$$
\Delta y_i(I_{k+1}, I_{k}) = f(I_{k+1}, \mathbf{c_i}) - f(I_{k}; \mathbf{c_i}) \quad ^\forall k = 1, \dots , K - 1
$$

Once treatment effects $\Delta y_i(I_{k+1}, I_{k})$ have been identified for every $k$, economic analysis can be conduced to identify the EOIR for site $i$. The profit change from increasing the input rate from $I_{1}$ to $I_{k}$ is

$$
P_c (\sum_1^{k-1} \Delta y_i(I_{k+1}, I_{k})) - P_I \times (I_k - I_0).
$$

The EOIR is the value $I_k$ that maximizes the profit gain "relative" to profits from $I_1$. ($I_1$ may be the EOIR if no other rates surpass the profit of $I_1$.)

@@

```{r ct-ex, fig.width=4, fig.height=4, fig.align="center", fig.cap = "An illustrative example of causal tree"}
# knitr::include_graphics("./GitControlled/Writing/Re_JournalSubmission/ctree_ex.pdf")
readPNG("./GitControlled/Writing/Re_JournalSubmission/ctree_ex.png") %>% grid.raster()
```

@@

## CF vs. Prediction-oriented ML methods

Whether the indirect (yield-prediction-first) and direct approach is used can lead to meaningful statistical differences when it comes to EOIR estimation. The conventional prediction-oriented ML methods are designed to predict yield, $y_i(I,\mathbf{c})$, well, but not necessarily treatment effects. Since the loss function used in prediction-oriented ML uses (typically squared) residuals in yield prediction, prediction-oriented ML is naturally optimized for predicting the "level" of yield instead of identifying the causal impact of the input. CF, on the other hand, is designed specifically to estimate the causal impact of input. Indeed, @Wager2018a used Monte Carlo simulations to show that CF provided more accurate estimates of heterogeneous treatment effects than those provided by RF. However, the underlying data generating process of their simulations was extremely simple, with the treatment effect influenced by only a single variable. CF has never been tested for more complicated data generating processes, such as those represented by yield response functions in which the treatment effect of an input can be affected by multiple variables in a non-linear fashion. Moreover, CF has never been compared to more advanced ML methods such as BRF and CNN that are considered to have prediction capabilities greater than RF's. The study reported here builds on @Wager2018a to compare CF to these prediction-oriented ML methods (not just RF, but also BRF and CNN) in more realistic agronomic experiment settings.

@@

Monte Carlo simulations were conducted for nitrogen experiments for corn production to estimate the economically optimal nitrogen rate (EONR). The first step in our simulation was to generate one thousand synthetic fields that "resembled" an actual field. <span style="color:red"> After generating one thousand synthetic fields, we repeated to evaluate each of the ML methods by one thousand times. In an individual simulation round, one of the fields was used as a training data set. As a validation data set, a different field from a training data set in that simulation round was used.</span> <!-- For each simulated field, we generated a training data set and a validation data set. --> Field characteristics and yield were simulated based on the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training data set, and checked the statistical performance of the predicted site-specific EONR using the validation data set. Detailed descriptions of these steps are provided below. The R codes that implement the simulation analyses are available at https://github.com/tmieno2/ML_VRA.git.

@@

The simulated field was created based on an actual 51-ha Illinois field on which the Data-Intensive Farm Management (DIFM) project [@bullock2019] has run multiple OFPEs. Figure \@ref(fig:field-map) illustrates that simulated trials consisted of $384$ <span style="color:red"> 18-m </span> $\times$ <span style="color:red"> 73-m </span> "plots," each of which was assigned an N fertilizer application rate. Each plot was made up of four (4-rows $\times$ 1-column) "subplots," which were unit of analysis used in subsequent statistical analysis. Each subplot contained of a 6-rows $\times$ 6-columns grid of thirty-six <span style="color:red"> 3-m </span> $\times$ <span style="color:red"> 3-m </span> "cells." All field characteristics and yield data were generated at the cell level.

```{r field-map, fig.width=4, fig.height=3, fig.cap = "Division of field: plots, subplots and cells"}
# knitr::include_graphics("./GitControlled/Writing/Re_JournalSubmission/field_devision_v2.png")
readPNG("./GitControlled/Writing/Re_JournalSubmission/field_devision_v2.png") %>% grid.raster()
```

@@

We modeled the yield response function as having the following Mitscherlich-Baule functional form:

\begin{equation}
f(N, \mathbf{c}) = ymax(\mathbf{c})(1-exp(\alpha(\mathbf{c}) + \beta(\mathbf{c}) N)) + \varepsilon. (\#eq:eqn3)
\end{equation}

The first term on the right-hand side of Eq. \@ref(eq:eqn3) is the deterministic yield component, which was generated assuming the commonly-used MB functional form, which is flexible in its shape, allows for the existence of a yield plateau, and has been shown to be a plausible specification for N fertilization models for corn [@frank1990; @llewelyn1997; @paris1992].
The variable $N$ in Eq. \@ref(eq:eqn3) represents the nitrogen fertilizer application rate (kg/ha) which the trial designs varied by plot. Also in Eq. \@ref(eq:eqn3), $\alpha$, $\beta$, and $ymax$ are functions of a vector of field characteristics variables $\mathbf{c}=(c_1,...,c_K)$, which vary spatially on the field. For example, $\alpha$ and $ymax$, which affect yield independently from the applied N fertilizer rate, might be interpreted as depending on the ingredient soil N content supplied by residual N, soil organic matter, etc. $\beta$ can be interpreted as reflecting how efficiently the crop uses N to create grain mass, and might depend on soil properties that affect the availability of applied N fertilizer to the crop, say due to N immobilization, leaching, denitrification, and volatilization [@johnson2005nitrogen; @alva2006nitrogen]. The error term of the yield $\varepsilon$ represents the composite of all the unobserved factors that affect crop yield beyond $\alpha$, $\beta$, $ymax$ and $N$.

A fictional "researcher" is modeled as having limited knowledge about which variables make up $\mathbf{c}$, or even the number of elements of $\mathbf{c}$, and has limited ability to measure the values that many of the variables that constitute $\mathbf{c}$ take on across the field. The researcher is also assumed to not know the functional forms of the yield response function.

### Cell-specific Yield Functions and Economically Optimal N Rates

Let $i \in \{1,2,...,1536\}$ index a simulated field's subplots and $j \in \{1,2,...,36\}$ index cells within a subplot. Let $N^{i}$ be the N fertilization rate on plot $i$, and let $\mathbf{c}^{i,j}$ be the value of $\mathbf{c}$ on cell $(i,j)$, Eq. \@ref(eq:eqn4) shows the yield level on cell $j$ in subplot $i$ is affected by the N rate applied on subplot $i$:

\begin{equation}
f(N^{i}, \mathbf{c}^{i,j}) = ymax(\mathbf{c}^{i,j})(1-exp(\alpha(\mathbf{c}^{i,j}) + \beta(\mathbf{c}^{i,j}) N^{i})) + \varepsilon^{i,j}. (\#eq:eqn4)
\end{equation}

@@

\begin{equation}
f^{i,j}(N) = ymax^{i,j}(1-exp(\alpha^{i,j} + \beta^{i,j} N^{i})) + \varepsilon^{i,j}. (\#eq:eqn5)
\end{equation}

Parameters $\alpha^{i,j}$, $\beta^{i,j}$, $\beta^{i,j}$, and $\varepsilon^{i,j}$ were modeled as spatially autocorrelated using unconditional Gaussian geostatistical simulation based on the spherical variogram model. The `gstat` package (version 2.0.6 [@gstat1; @gstat2]) in R was used for this purpose. Table \@ref(tab:variogram-parameters) shows the main variogram parameter values used in generating $\alpha^{i,j}$, $\beta^{i,j}$, $ymax^{i,j}$, and $\varepsilon^{i,j}$, which were chosen to generate yields consistent with those observed in experiments conducted by the DIFM project (Codes used to generate parameter values are available at https://github.com/tmieno2/ML_VRA.git.) Figure \@ref(fig:fig-parameters) shows an example of the generated parameters.

@@

```{r variogram-parameters, tab.cap = "The parameters for the variogram modeling"}
variogram_tb
```

```{r fig-parameters, fig.width=6, fig.height=5, fig.cap = "An illustrative example of spatial distributions of field characteristics"}
# field_alpha + field_beta + field_ymax
grid.arrange(field_ymax, field_alpha, field_beta, field_m_error, ncol=2, nrow=2)
```

@@

\begin{equation}
N_{opt}^{i,j} = \operatorname*{argmax}_{N}(P_{C} f^{i,j}(N) - P_N N), (\#eq:eqn6)
\end{equation}

>>>>

---
title: "Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation"
author: Shunkei Kakimoto^[Corresponding author, Department of Agricultural Economics, University of Nebraska-Lincoln 102 Filley Hall 1625 Arbor Drive Lincoln, NE 68583, USA, E-mail address; skakimoto3@hunskers.unl.edu], Taro Mieno^[Department of Agricultural Economics, University of Nebraska-Lincoln, Lincoln, NE 68583, USA], Takashi S. T. Tanaka^[Artificial Intelligence Advanced Research Center, Faculty of Applied Biological Sciences, Gifu University, Gifu 501-1193, Japan], and David S. Bullock^[Department of Agricultural and Consumer Economics, University of Illinois, Urbana, IL 61801, USA]
output:
  bookdown::word_document2:
    fig_caption: yes
    number_sections: yes
    global_numbering: true
    # reference_docx: word_template.docx
    reference_docx: word_template_format.docx
    # pandoc_args: ["-Fpandoc-crossref"]
bibliography: ML_VRA.bib
csl: computers-and-electronics-in-agriculture.csl
eqnPrefix: 
      - "Eq"
abstract: "Estimating site-specific crop yield response to changes in input (e.g., seed, fertilizer) management is a critical step in making economically optimal site-specific input recommendations. Past studies have attempted to estimate yield response functions using various Machine Learning (ML) methods, including the Random Forest (RF), Boosted Random Forest (BRF), and Convolutional Neural Network (CNN) methods. This study proposes use of the Causal Forest (CF) model, which is one of the emerging ML methods that comprise \"Causal Machine Learning.\" Unlike previous yield-prediction-oriented ML methods, CF focuses strictly on estimating heterogeneous treatment effects (changes in yields that result from changes in input application rates) of inputs. We report results of using Monte Carlo simulations assuming various production scenarios to test the effectiveness of CF in estimating site-specific economically optimal nitrogen rates (EONRs), comparing CF with the yield-prediction-oriented ML methods RF, BRF, and CNN. CF's estimations of site-specific EONRs were superior under all scenarios considered. Furthermore, we also show that the quality of a model's yield predictions provides little if any information about the quality of its EONR predictions."  
---

@@

# opts_knit$set(root.dir = here())
opts_knit$set(root.dir = "~/Dropbox/ResearchProject/CF_for_VRA")

@@

```{r, cache = FALSE}
# === packages ===#
library(tidyverse)
library(data.table)

library(png)
library(RColorBrewer)
library(patchwork)
library(plotly)
library(magrittr)
library(ggplot2)
library(viridis)
library(tmap)
library(ggthemes)
library(mgcv)
library(ggpubr)
library(flextable)
library(officer)
library(officedown)
library(modelsummary)
library(raster)
library(bookdown)
library(png)
library(grid)
```

```{r analyses, results = "hide", fig.show='hide', cache = FALSE}
#/*=================================================*/
#' # Preparation
#/*=================================================*/
# This is where you prepare data sets and results 
# library(here)
# setwd(here())
#/*----------------------------------*/
#' ## Create figures and tablesb
#/*----------------------------------*/
# --- Convert an Rmd file into a pure R file using purl() ---#
knitr::purl("./GitControlled/Codes/PrepareResults_1.Rmd", output = "./GitControlled/Codes/PrepareResults_1.R")

# --- Read the simulation results and prepare figures and tables ---#
source("./GitControlled/Codes/PrepareResults_1.R")
```

@@

Information about crop yield response input management is critical for the identification of site-specific economically optimal input rates (EOIRs) [@bullock1998does], and the difficulty in estimating site-specific yield response functions accurately within a field is the primary explanation of the persistent relatively low the adoption rate of variable rate crop input application technology [@lowenberg2019setting]. Historically, site-specific yield response functions have been estimated using the data generated from strip trials conducted on only parts of fields [@makowski2001statistical; @scharf2002corn; @bullock2002adding; @mamo2003spatial; @hurley2004estimating; @anselin2004spatial; @scharf2005field; @ruffo2006site; @Miao2006; @puntel2019development].
But recently whole-field on-farm precision experimentation (OFPE) [@bullock2019] has allowed implementation of improved statistical experimental designs (e.g., Latin Square, Randomized Block) on entire fields [@laurent_kyveryga_makowski_miguez_2019; @licht_witt_2019]. Data generated from whole-field OFPEs typically have greater local variation in experimental input application rates and soil/field characteristics than do strip trial designs, which has improved the statistical identification of heterogeneous impacts of input management changes on yields. The resultant availability of more spatially detailed experiment data has encouraged a resurgence of interest in estimating site-specific yield response functions. For example, @barbosa2020modeling, @krause2020random and @Gardner21 used OFPE data to estimate site-specific yield response functions and economically optimal site-specific N fertilizer and/or seed rates. These recent studies used machine learning (ML) methods extensively in place of more traditional parametric modeling approaches. @barbosa2020modeling applied the Convolutional Neural Network approach, @krause2020random used Random Forest-based approaches, and @Gardner21 used the shape-constrained generalized additive model. 
But this trend of applying ML methods may raises concerns. Historically, agronomists have worked meticulously to develop clever experimental designs to identify causes of yield response to input management. But the aforementioned recent applications of ML methodologies pay little attention to identify such causes, instead attempting to validate their models by estimating yield prediction accuracy. But the ultimate goal of OFPE is the accurate prediction of EOIRs, not yields. These are distinct objectives, and achievement of one need not imply the other.

The lack of and need for causal interpretation of the results of ML research applications have been increasingly recognized, in particular in economics and medicine [@athey2017beyond; @athey2018impact; @Athey2019a; @moraffah2020causal; @Storm2020; @arti2020research; @scholkopf2021toward], and causal machine learning (CML) methods were recently developed in response. Unlike traditional prediction-oriented ML methods, CML focuses on identifying causal impacts of an event (in our context, of an increase or decrease in the input application rate). The growing academic literature applying CML methods includes @richens2020improving, who's medical research pointed out that the inability of the traditional ML methods to identify cause-and-effect relationships between disease and patients' characteristics could lead to misdiagnoses, and showed that CML-based diagnoses could improve the accuracy of diagnostic prediction. In addition, @carbo2020machine used a CML to identify characteristics causing bank customers to adopt online banking services, and @bozorgi2020process applied CML to study the mining processes.

In this paper, we use the Causal Forest (CF) approach developed by @athey2016recursive and @Wager2018a to examine the accuracy of methods that estimate site-specific economically optimal N fertilizer rates (EONRs). We compare the EONR prediction performances of our CF-based approach to the widely used ML methods Random Forest (RF), Boosted Random Forest (BRF), and Convolutional Neural Network (CNN), and we examine how yield prediction performance is related to EONR prediction performance, which allow us to examine whether model selection should be based on yield prediction accuracy when the ultimate goal is EOIR identifying. We use Monte Carlo simulations in which we synthetically generate yield and soil/field characteristics datasets to create stylized "field." Because real-world "true" EOIRs are unobservable, real-world data cannot be used to test the accuracy of statistical methods estimating EOIRs. This important difference contrasts starkly with yield prediction, since actual yields can be ground-truthed.

# Background:Using Causal Forest Methods to Predict Changes in Yield

In the following, we use basic microeconomic theory to describe the fundamental information necessary to identify EOIRs in general: the site-specific treatment effect of input application on yield. We then discuss the causal forest method in detail, show how it is particularly suitable to obtain this fundamental information, and contrast it to more traditional prediction-oriented ML methods, such as RF, BRF, and CNN.

@@

Elementary calculus shows that the necessary condition for maximization is, that $\partial f(I, \mathbf{c})/\partial I - w =0$, which implicitly defines the economically optimal input application rate as a function $I^{*}$ of prices and soil/field characteristics, and may be arranged as \@ref(eq:eqn2):

@@

Identity \@ref(eq:eqn2) states that the necessary elements of the information needed for profit maximization are the input price, the output price, and the partial derivative of the yield function with respect to the input. While knowledge of the yield function $f(I, \mathbf{c})$ in Eq. \@ref(eq:eqn1) is sufficient to solve the maximization problem above (since its derivative can be calculated), identity \@ref(eq:eqn2) shows that it is not a necessary condition; rather, knowing the function's partial derivative is necessary.

In the following, we discuss a method of using Causal Forest methods that can provide statistical advantages to the estimation of economically optimal input application rates because it is built to estimate approximations of the derivative $\partial f/\partial I$ directly, and not the yield function $f$ itself.

@@

Recent studies using ML methods to estimate the effectiveness of various agricultural input management strategies have based their analyses on estimates of a yield function, $f(I, \mathbf{c})$ (e.g., @krause2020random; @barbosa2020modeling; @Gardner21). In the following, we present a Causal Forest methodology that does not rely on estimation of how the level of yield depends on the input application rate $I$, but rather relies on a discrete version of Eq. \@ref(eq:eqn2), examining the discrete changes in yield that result from discrete *changes* in the input application rate $I$.

Consider an increasing ordering of input application rates $I_1, I_2,...,I_K$. Among these K rates, for a generic site $i$ a specific EOIR can be identified if the *treatment effect* function $\Delta y_i(I_{k+1}, I_{k})$ defined in Eq. \@ref(eq:eqn3)is known. This function is defined as the discrete change in yield that results from a discrete change in the input level from $I_k$ to $I_{k+1}$, and it depends on the characteristics at the site analyzed:

\begin{equation}
\Delta y_i(I_{k+1}, I_{k}) = f(I_{k+1}, \mathbf{c_i}) - f(I_{k}; \mathbf{c_i}) \quad ^\forall k = 1, \dots , K - 1. (\#eq:eqn3)
\end{equation}

Once treatment effects $\Delta y_i(I_{k+1}, I_{k})$ have been identified for every $k$, economic analysis can be conducted to identify the EOIR for site $i$. The profit change from increasing the input rate from $I_{1}$ to $I_{k}$ is

\begin{equation}
p (\sum_1^{k-1} \Delta y_i(I_{k+1}, I_{k})) - w \times (I_k - I_0). (\#eq:eqn4)
\end{equation}

The EOIR is the value $I_k$ that maximizes the profit gain "relative" to profits from $I_1$. ($I_1$ may be the EOIR if no other rates generates greater profit.)

@@

```{r ct-ex, out.width="50%", fig.align="center", fig.cap = "An illustrative example of causal tree"}
# knitr::include_graphics("./GitControlled/Writing/Re_JournalSubmission/ctree_ex.pdf")
readPNG("./GitControlled/Writing/Re_JournalSubmission/ctree_ex.png") %>% grid.raster()
```

@@

## Advantages of Causal Forest Analysis in the Context of On-farm Precision Experimentation

Whether the indirect (yield-prediction-first) or direct approach is used can lead to meaningful statistical differences in EOIR estimation. The conventional prediction-oriented ML methods are designed to predict yield, $y_{i}(I) \equiv f(I, c_{i})$, well, but not necessarily treatment effects. Since the loss function used in prediction-oriented ML uses (typically squared) residuals in yield prediction, prediction-oriented ML is naturally optimized for prediction the "level" of yield instead of identifying the causal impact of the input. But CF is designed specifically to estimate the casual impact of the input. Indeed, @Wager2018a used Monte Carlo simulations to show that CF provided more accurate estimates of heterogeneous treatment effects than those provided by RF. However, the underlying data generating process of their simulations was extremely simple, with the treatment effect influenced by only a single variable. CF has never been tested for more complicated data generating processes, such as those represented by yield response functions in which the treatment effect of an input can be affected by multiple variables in a non-linear fashion. Moreover, CF has never been compared to more advanced ML methods such as BRF and CNN that are considered to have prediction capabilities greater than RF's. Our report builds on @Wager2018a to compare CF to these prediction-oriented ML methods (not just RF, but also BRF and CNN) in more realistic agronomic experiment settings.

@@

Monte Carlo simulations were conducted for nitrogen experiments for corn production to estimate site-specific EONRs. The simulations’ first step generated one thousand synthetic fields that "resembled" an actual field. Then, for each simulation round, one of the fields was used as a training data set, and another field that is different from the one used as a training data set in that round was used as a validation data set. Field characteristics and yield were simulated assuming the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training data set, and checked the statistical performance of the predicted site-specific EONR using the validation data set. Detailed descriptions of these steps are provided below. The *R* and *python* codes that implement the simulation analyses are available at https://github.com/tmieno2/ML_VRA.git.

@@

The simulated field was created based on an actual 51-ha Illinois field on which the Data-Intensive Farm Management (DIFM) project [@bullock2019] has run multiple OFPEs. Figure \@ref(fig:field-map) illustrates that simulated trials consisted of $384$ 18.3 m $\times$ 73.2 m "plots," each of which was assigned an N fertilizer application rate. Each plot was made up of four (4-rows $\times$ 1-column) 18.3 m $\times$ 18.3 m "subplots," which were unit of analysis used in subsequent statistical analysis. Each subplot contained of a 6-rows $\times$ 6-columns grid of thirty-six 3.05 m $\times$ 3.05 m "cells." All field characteristics and yield data were generated at the cell level.

```{r field-map, out.width="50%", fig.cap = "Partitioning of field into plots, subplots and cells"}
# knitr::include_graphics("./GitControlled/Writing/Re_JournalSubmission/field_devision_v2.png")
readPNG("./GitControlled/Writing/Re_JournalSubmission/field_devision_v2.png") %>% grid.raster()
```

@@

We modeled the yield response function as having the following Mitscherlich-Baule functional form, which is flexible in its shape, allows for the existence of a yield plateau, and has been shown to be a plausible specification for N fertilization models for corn [@frank1990; @llewelyn1997; @paris1992].

\begin{equation}
f(N, \mathbf{c}) = ymax(\mathbf{c})(1-exp(\alpha(\mathbf{c}) + \beta(\mathbf{c}) N)) + \varepsilon. (\#eq:eqn5)
\end{equation}

The first term on the right-hand side of Eq. \@ref(eq:eqn5) is the deterministic yield component. The variable $N$ represents the nitrogen fertilizer application rate (kg/ha) which the trial designs varied by plot. Also in Eq. \@ref(eq:eqn3), $\alpha$, $\beta$, and $ymax$ are functions of a vector of field characteristics variables $\mathbf{c}=(c_1,...,c_K)$, which vary spatially on the field. For example, $\alpha$ and $ymax$, which affect yield independently from the applied N fertilizer rate, might be interpreted as depending on the ingredient soil N content supplied by residual N, soil organic matter, etc. $\beta$ can be interpreted as reflecting how efficiently the crop uses N to create grain mass, and might depend on soil properties that affect the availability of applied N fertilizer to the crop, say due to N immobilization, leaching, denitrification, and volatilization [@johnson2005nitrogen; @alva2006nitrogen]. The error term of the yield $\varepsilon$ represents the composite of all the unobserved factors that affect crop yield beyond $\alpha$, $\beta$, $ymax$ and $N$.

A fictional "researcher" is modeled as not knowing the function form of $f$, as having limited knowledge about which variables make up $\mathbf{c}$, or even the number of elements of $\mathbf{c}$, and as having limited ability to measure the values that many of the variables that constitute $\mathbf{c}$ take on across the field.

### Cell-specific Yield Functions and True Economically Optimal N Rates

Let $i \in \{1,2,...,1536\}$ index a simulated field's subplots and $j \in \{1,2,...,36\}$ index cells within subplots. Let $N^{i}$ be the N fertilization rate on plot $i$, and let $\mathbf{c}^{i,j}$ be the value of $\mathbf{c}$ on cell $(i,j)$, Eq. \@ref(eq:eqn6) shows the yield level on cell $j$ in subplot $i$ is affected by the N rate applied on subplot $i$:

\begin{equation}
f(N^{i}, \mathbf{c}^{i,j}) = ymax(\mathbf{c}^{i,j})(1-exp(\alpha(\mathbf{c}^{i,j}) + \beta(\mathbf{c}^{i,j}) N^{i})) + \varepsilon^{i,j}. (\#eq:eqn6)
\end{equation}

@@

\begin{equation}
f^{i,j}(N) = ymax^{i,j}(1-exp(\alpha^{i,j} + \beta^{i,j} N^{i})) + \varepsilon^{i,j}. (\#eq:eqn7)
\end{equation}

Parameters $\alpha^{i,j}$, $\beta^{i,j}$, $\beta^{i,j}$, and $\varepsilon^{i,j}$ were modeled as spatially autocorrelated using unconditional Gaussian geostatistical simulation based on the spherical variogram model. The `gstat` package (version 2.0.6 [@gstat1; @gstat2]) in R was used for this purpose. Table \@ref(tab:variogram-parameters) shows the main variogram parameter values used in generating $\alpha^{i,j}$, $\beta^{i,j}$, $ymax^{i,j}$, and $\varepsilon^{i,j}$, which were chosen to generate yields consistent with those observed in experiments conducted by the DIFM project. (Codes used to generate parameter values are available at https://github.com/tmieno2/ML_VRA.git.) The range and nugget of all parameters were assumed to be $400$ m and 0. The means of $ymax$, $\alpha$, $\beta$, and $\varepsilon$ were assumed to be $12000$ kg/ha, $-0.5$, $0$ and $0$, and their sills were assumed to be $2.0 \times 10^{6}$ kg/ha, $0.02$, $1.0$ and $0.015$. Figure \@ref(fig:fig-parameters) shows an example of the generated parameters.

@@

```{r fig-parameters, fig.width=6, fig.height=5, fig.cap = "An Illustrative Example of Spatial Distributions of Field Characteristics"}
# field_alpha + field_beta + field_ymax
grid.arrange(field_ymax, field_alpha, field_beta, field_m_error, ncol=2, nrow=2)
```

@@

\begin{equation}
N_{opt}^{i,j} = \operatorname*{argmax}_{N}(P_{C} f^{i,j}(N) - P_N N), (\#eq:eqn8)
\end{equation}

where, $p =\$0.138/kg$ and $w =\$1.323/kg$ were the assumed prices of the corn and N fertilizer. Figure \@ref(fig:fig-optN) maps an example of the identified "true" cell-specific optimal N rates.
